% This document will act as a rough timeline for the thesis projection

\documentclass[12pt, letterpaper]{article}
\usepackage{syntonly}
% Next three packages are on the off chance that graphs or mathematics are needed here.
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}

%\syntaxonly % Use this when adding new lines to check for compile errors.
%              (increases run speed and prevents document from being produced.)

\begin{document}
  \section{Timeline}
  \begin{itemize}
    \item[$\cdot$] September:
    \begin{itemize}
      \item[-] Discuss starting process, goals, and methodology for the project.
      \item[-] Start collecting data and begin code work to implement methodology
               (i.e., begin analysis).\\
    \end{itemize}
    
    \item[$\cdot$] October:
    \begin{itemize}
      \item[-] Continue code work and analysis.
      \item[-] Likely more discussions over methods and perhaps some preliminary results.\\
    \end{itemize}
    
    \item[$\cdot$] November:
    \begin{itemize}
      \item[-] Continue code work and analysis.
      \item[-] Begin creating poster with current results for presentation.\\
    \end{itemize}
    
    \item[$\cdot$] December:
    \begin{itemize}
      \item[-] Continue code work and analysis some more.
      \item[-] Continue working on poster and associated presentation (begin to finalize
               it here would be ideal).\\
    \end{itemize}
    
    \item[$\cdot$] January:
    \begin{itemize}
      \item[-] Finish poster if not finished already.
      \item[-] Present poster with some preliminary results.
      \item[-] Begin writing out thesis
      \begin{itemize}
        \item[$\rightarrow$] Should be enough at this point for Chapters 1 - 4 in the
                             outline and some of Chapter 5.\\
      \end{itemize}
    \end{itemize}
    
    \item[$\cdot$] February:
    \begin{itemize}
      \item[-] More code work and analysis.
      \item[-] Continue to work on thesis.
      \item[-] Consider starting presentation for seminar here.\\
    \end{itemize}
    
    \item[$\cdot$] March:
    \begin{itemize}
      \item[-] More analysis. For the thesis, this might start wrapping up soon.
      \item[-] Continue to work on thesis.
      \item[-] Start working on seminar presentation if not already. Otherwise, continue
               working on it.\\
    \end{itemize}
    
    \item[$\cdot$] April:
    \begin{itemize}
      \item[-] Continue analysis.
      \item[-] Continue working on thesis.
      \item[-] Finalize and give seminar presentation.
      \begin{itemize}
        \item[$\rightarrow$] At this point, it may be worth shifting the main focus from
                             data analysis to working on the presentation and thesis.\\
      \end{itemize}
    \end{itemize}
    
    \item[$\cdot$] May:
    \begin{itemize}
      \item[-] Start wrapping up data analysis if not already.
      \item[-] Continue working on/editing thesis.
      \begin{itemize}
        \item[$\rightarrow$] A workable draft (or at least a rough draft) of the thesis
                             should be done by this point.\\
      \end{itemize}
    \end{itemize}
    
    \item[$\cdot$] June:
    \begin{itemize}
      \item[-] Ensure data analysis is done by this point.
      \item[-] Edit thesis and start to prepare for thesis defense.\\
    \end{itemize}
    
    \item[$\cdot$] July:
    \begin{itemize}
      \item[-] Finish thesis and continue preparing for thesis defense.
      \item[-] Defend thesis.
      \item[-] Post defense, begin month long sleep/hibernation.\\
    \end{itemize}
    
    \item[$\cdot$] August:
    \begin{itemize}
      \item[-] Continue month of continuous sleep/hibernation until new semester.
      \item[-] Repeat the whole process again with a much larger project and longer time
               period and more publications for PhD dissertation.\\
    \end{itemize}
  \end{itemize}
  
  A few notes to be had here:
  \begin{itemize}
    \item[-] Readings are assumed to be seeded throughout the whole processes and are
             omitted for the sake of brevity.
    \item[-] Meetings and discussions are also assumed to happen throughout the process
             and are also omitted after October for the sake of brevity.
    \item[-] This point was forgotten before written...
  \end{itemize}
\end{document}
      